---
title: "Managing Many Models"
subtitle: "R Programming Practice with Hardley Wickham"
author: "James Mba Azam"
date: "April 2, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("gapminder")
library("tidyverse")
```
###INTRODUCTION
1. Convert models to tidy dataframes with the BROOM package
2. Use DPLYR for working with dataframes
3. Use PURRR for lists
4. Use TIDYR for converting between dataframe and lists of dataframes
```{r}
#Import data from the gapminder package (very cool package with great data)
# year1950 is a column that contains the number of years since 1950
gapminder <- gapminder %>% mutate(year1950 = year-1950)


#nested data

by_country <- gapminder%>%
  group_by(country,continent) %>%
  nest()

#country_model() fits each country data to a linear model

country_model <- function(df){
  lm(lifeExp ~ year1950, data = df)
}


models <- by_country %>%
  mutate(model=map(data, country_model))


models %>% filter(continent=="Africa")
```
what can we do with a list of linear models? not very much

we can convert our data in tidy data using the broom package

what sort of data can we get from our models?
In BROOM, `glance` gives the model summaries,
`tidy()` gives the estimates, and `augment` gives the stats per observation

```{r}
library("broom")

models <- models %>%
  mutate(
    glance = map(model,broom::glance),
    rsq =glance %>%map_dbl("r.squared"),
    tidy = map(model, broom::tidy),
    augment = map(model, broom::augment)
  )
models
unnest(models,data)
unnest(models,glance, .drop = TRUE)# %>%View()
unnest(models,rsq)#%>%View()
unnest(models,tidy)#%>%View()
unnest(models,augment)#%>%View()
```


### CONCLUSION
1. Store related objects in list-clumns 
2. Learn functional programming to concentrate on the verb 
and not the object
3. Use broom to convert models to tidy data



