---
output:
  html_document: default
  number_sections: yes
  pdf_document: default
---
<div class="row">

<div class="col-lg-3"></div>

<div class="col-lg-6">
<br>R Programming Practice</br>
<br>Authored By: James Mba Azam</br>
<br>Last Updated: `r Sys.Date()`</br>
</div>

<div class="col-lg-3"></div>

</div>
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("gapminder")
library("tidyverse")
```
# MANAGING MANY MODELS (By HARDLEY WICKHAM)
### Summary
1. Convert models to tidy dataframes with the BROOM package
2. Use `dplyr` for manipulating dataframes
3. Use `purrr` for list-wise operations
4. Use `tidyr` for converting between dataframe and lists of dataframes
```{r}
#Import data from the gapminder package (very cool package with great data)
# year1950 is a column that contains the number of years since 1950
gapminder <- gapminder %>% mutate(year1950 = year-1950)


#nested data

by_country <- gapminder%>%
  group_by(country,continent) %>%
  nest()

#country_model() fits each country data to a linear model

country_model <- function(df){
  lm(lifeExp ~ year1950, data = df)
}


models <- by_country %>%
  mutate(model=map(data, country_model))


models %>% filter(continent=="Africa")
```
what can we do with a list of linear models? not very much

we can convert our data in tidy data using the broom package

what sort of data can we get from our models?
In BROOM, `glance` gives the model summaries,
`tidy()` gives the estimates, and `augment` gives the stats per observation

```{r}
library("broom")

models <- models %>%
  mutate(
    glance = map(model,broom::glance),
    rsq =glance %>%map_dbl("r.squared"),
    tidy = map(model, broom::tidy),
    augment = map(model, broom::augment)
  )
models
unnest(models,data)
unnest(models,glance, .drop = TRUE)# %>%View()
unnest(models,rsq)#%>%View()
unnest(models,tidy)#%>%View()
unnest(models,augment)#%>%View()
```


### Conclusion
1. Store related objects in list-clumns 
2. Learn functional programming to concentrate on the verb 
and not the object
3. Use broom to convert models to tidy data


# DATA VISUALIZATION
```{r}
library("ggplot2")
library("dplyr")
library("reshape2")
library("lubridate")
library("dplyr")
library("tidyr")
library("ggplot2")
library("scales")
library("gridExtra")
```

```{r}
data("economics", package = "ggplot2")
head(economics)
# Create the plot
ggplot(data = economics) + geom_line(aes(x = date, y = unemploy))
```
##Scatter Plots
It’s primarily used by people doing analysis and not reporting. The inventor of the scatter plot was looking for the correlation between area and population.
```{r}
data("midwest", package = "ggplot2")
ggplot(data=midwest, aes(x=area, y=poptotal))+geom_point()


#further customization for illustration purposes
ggplot(data=midwest, aes(x=area, y=poptotal, col=state, size=popdensity))+geom_point()


#labelling of the axis
ggplot(data=midwest, aes(x=area, y=poptotal))+geom_point(aes(col=state, size=popdensity))+
  geom_smooth(method="loess", se=F) + 
  xlim(c(0, 0.1)) + 
  ylim(c(0, 500000)) + 
  labs(subtitle="Area Vs Population", 
       y="Population", 
       x="Area", 
       title="Scatterplot", 
       caption = "Source: midwest (ggplot2)")
#note that the se=F is to indicate that error bars should not be shown
```

##Density Plots
Density plots are similar to histograms except that they illustrate percentages. They are used to display the distribution of data.

we will be using the R-base mpg data. Let us preview the data now.
```{r}
head(mpg)
img <- ggplot(data = mpg, aes(cty))
img + geom_density(aes(fill=factor(cyl)), alpha=0.8) + 
    labs(title="Density plot", 
         subtitle="City Mileage Grouped by Number of cylinders",
         caption="Source: mpg (R-base)",
         x="City Mileage",
         fill="# Cylinders")
```

By using the assignment feature of `ggplot()`. We can keep the assigned object `img`. It is very useful if we want to use `img` as the fundamental plot on which to explore other visualizations. It is also handy if you need to save the output of a chart to an image file.

By simply writing `aes(cty)`, `ggplot` assumes we mean `aes(x = cty)`. 
`alpha` is the transparency (or opacity) of the area under the curve.

As mentioned earlier, we can explore other plots by easily plotting on `img` since it has the basic layer established on it. For example, we can explore the `class` variable as follows:
```{r}
img + geom_density(aes(fill=factor(class)), alpha=0.8) + 
    labs(title="Density plot", 
         subtitle="City Mileage Grouped by Class",
         caption="Source: mpg (R-Base)",
         x="City Mileage",
         fill="Class")
```
##Histograms
 We are going to show the city mileage with a histogram
 
```{r}
img + geom_histogram(bins = 20,aes(col="red"))+
  labs(title="Histogram", 
         caption="Source: mpg",
         x="City Mileage")
```
#Bar/Column Chart
Technically, the term “column chart” can be used when the bars run vertically.

We need to first prepare the data. We shall use the `table` function to create a frequency table of all the car manufacturers. `ggplot` works best with data frames and so, we will convert the frequency table to a data frame as well. All this can be achieved as follows
```{r}
dat <- as.data.frame(table(mpg$manufacturer))
#set a theme
theme_set(theme_bw())

img <- ggplot(data=dat, aes(x=Var1, y=Freq))
img + geom_bar(stat = "identity", width = 0.5, fill="tomato2") +
    labs(title="Bar Chart", subtitle="Manufacturer of vehicles", 
           caption="Source: Frequency of Manufacturers from 'mpg' dataset") +
      theme(axis.text.x = element_text(angle=65, vjust=0.6))

```

<b>Side Notes:</b>
<ol>
<li>With `theme_set()`, you can set a background theme for your graphs. There are several default ones you can choose from. Just type `theme_` and press the `tab` button.</li>
<li>`theme(axis.text.x = element_text(angle=65, vjust=0.6))` tilts the text along that angle on the x-axis labelling. </li>
<li>`vjust=0.6` sets the distance between the text and the x-axis line.
<li> `stat="identity"` tells ggplot to use the actual value of `Freq`.</li>
</ol>

Column charts can be flipped around with the `coord_flip()` function.
```{r}
img <- ggplot(dat,aes(reorder(Var1,Freq),Freq))
img + geom_bar(stat = "identity",width = 0.5,fill="tomato2")+
    labs(title="Bar Chart", 
         subtitle="Manufacturer of vehicles", 
           x = 'Manufacturer',
           caption="Source: Frequency of Manufacturers from 'mpg' dataset") + theme(axis.text.x = element_text(angle=65, vjust=0.6)) + coord_flip()
```
Nest, we want to illustrate how `hwy` is related to the `manufacturer` and in terms `cyl`


```{r}
img <-  ggplot(mpg,aes(x=manufacturer,y=hwy,col=factor(cyl),fill=factor(cyl)))
img + geom_bar(stat='identity') + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6))
```
By default, the bars are stacked as shown above. In order to place them side by side for better comparison, we can set `position="dodge"` in the `geom_bar` properties. It will look like the following:
```{r}
img <-  ggplot(mpg,aes(x=factor(cyl),y=hwy,col=factor(cyl),fill=factor(cyl)))
img + geom_bar(stat='identity', position = "dodge") + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6))
```

Unless you are very patient to squeeze your eyes in order to read the graphs, it is obvious that the graph above is very difficult to read. One nice feature of `ggplot` is the `facet_wrap` function. It splits the graphs according to a variable you specify. Let us see how it works in the code below:
```{r}
img <-  ggplot(mpg,aes(x=factor(cyl),y=hwy,col=factor(cyl),fill=factor(cyl)))
img + geom_bar(stat='identity', position = "dodge") + 
  facet_wrap(~manufacturer)
```


##Spatial Maps
```{r  messages=FALSE, cache=FALSE}
library(ggmap)
library(ggalt)

# Get the Map ----------------------------------------------
colo_map <- qmap("Stellenbosch, South Africa",zoom = 7, source = "google")   

# Get Coordinates for Places ---------------------
colo_places <- c("Idas Valley",
                    "Stellenbosch University")

places_loc <- geocode(colo_places)  # get longitudes and latitudes


# Plot Open Street Map -------------------------------------
colo_map + geom_point(aes(x=lon, y=lat),
                             data = places_loc, 
                             alpha = 0.7, 
                             size = 2, 
                             color = "tomato") + 
                  geom_encircle(aes(x=lon, y=lat),
                                data = places_loc, size = 1, color = "blue")
```

